%% bare_jrnl.tex
% Sams
\documentclass[letterpaper,10pt]{IEEEtranTCOM}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl}
\usepackage{float}
% For reading with less pages wasted uncomment the next line
%\usepackage[margin=0.5in, paperwidth=8.5in, paperheight=11 in]{geometry}
%\usepackage[margin=0.5in]{geometry}
% For doublespacing uncomment the next 2 lines
\usepackage{setspace}   
%\doublespacing   
%\floatstyle{boxed} 
\floatstyle{ruled} 
\restylefloat{figure}
\normalsize
\hyphenation{op-tical net-works semi-conduc-tor}


\begin{document}
\title{Survey of Deployed Methods\\for Increasing Internet Address Space}

\author{William~Chimiak,~\IEEEmembership{Senior Member,~IEEE,}\\
    Samuel~Patton, \\and~Stephen~Janansky}

%\author{William~Chimiak,~\IEEEmembership{Senior Member,~IEEE,}\\ Samuel~Patton,\\ and~Stephen~Janansky}


\date{December 14, 2011}

% The paper headers
%\markboth{IEEE Computer}%
%{Submitted paper}

% make the title area
\maketitle

\begin{abstract}
%\begin{figure*}
%\end{figure*}
This paper provides the background for an extension to IPv4 called 
Enhanced IP (EnIP).  EnIP extends the address space of IPv4 in 
response to the IPv4 address depletion problem. This paper serves as 
a literature survey on IPv4, IPv6, and Carrier Grade NATs.  This survey was
used to inform research on Enhanced IP, the details of which are 
covered in a seperate paper.

\end{abstract}


% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IPv4 address depletion, IPv6, Enhanced IP.
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%% \IEEEpeerreviewmaketitle

\section{Introduction}

The rapid increase of mobile devices has created a large demand for
IP addresses.  At present, there are approximately 7 billion people and 3 billion usable IPv4 addresses.
It would be useful to have a unique address for each mobile device so that any device could call another.  
Presently this is not possible.  Since 1999, IPv6 has presented a similar
argument that a longer address would make VoIP applications simpler to deploy.  We analyzed IPv6 to learn
how it works.

In addition to IPv6 analysis, we also studied transitions during the life of IPv4.  On at least four occasions, IPv4
evolved to remedy address space exhaustion issues.  Further, we looked into the history of ARPANET
at a fundamental transition in its history: the switchover from NCP to TCP/IP.  Finally, we
performed a literature survey on Carrier Grade NAT technology.  Carrier Grade NAT can offer an immediate fix
to ISPs with limited IP address space.  Deployment of Carrier Grade NAT appears to have some additional costs in terms
of additional expense involved in law enforcement monitoring.  CGNs may also ossify the network to make future deployment
of IPv6 more difficult.

\section{Outline}
Section III of the paper discusses several transitions in the Internet Protocol
suite.  These transitions include NCP to TCP/IP, the evolution of IPv4, and IPv6.
The final section includes closing remarks.

\input{transitions}  %sam

%\include makes a pagebreak, input does not
%\include{conclusion}
\input{conclusion}

\bibliographystyle{plain}
\bibliography{ourrefs,rfc}

\end{document}
