%\documentclass[letterpaper,10pt]{article}
\documentclass[letterpaper,final,peerreview,10pt,onecolumn]{IEEEtranTCOM}

\usepackage{url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl}
\usepackage{float}
% For reading with less pages wasted uncomment the next line
% 2012 Jan 27 commented out as per editor request
%\usepackage[margin=0.7in, paperwidth=8.5in, paperheight=11in]{geometry}
% For doublespacing uncomment the next 2 lines
\usepackage{setspace}
%\doublespacing
%\floatstyle{boxed} 
\floatstyle{ruled} 
\restylefloat{figure}
\normalsize
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

% Title can use linebreaks \\ within to get better formatting as desired
\title{Enhanced IP}
\date{April 17, 2012}
\author{William~Chimiak,~\IEEEmembership{Senior Member,~IEEE,}\\ Samuel~Patton,\\ and~Stephen~Janansky}
%\author{William~Chimiak, Stephen~Janansky, Samuel~Patton }

% The paper headers
%\markboth{IEEE Computer}%
%{Submitted paper}

% make the title area
\maketitle


\begin{abstract}
This paper introduces an extension to IPv4 called Enhanced IP, or EnIP for
short.  EnIP offers a solution to IPv4 address depletion. EnIP
does not replace IPv4 but builds on top of it, maximizing backward
compatibility.  EnIP is currently deployed between two nodes at the University
of Maryland and the University of Delaware, connected via Internet 2.
Each network has an IPv4 gateway with hosts behind the gateway.
Packets between the two networks are routed over IPv4 but using addresses
of the following form: 65.127.221.1.10.1.1.1.  The first four bytes
is the gateway address and the second four bytes, carried in an IP option, is
the address of a host behind the gateway.  
  
Applications such as http, samba, and ssh have been demonstrated between 
the two EnIP networks \textbf{without modification} of the software or routers 
in the path.
\end{abstract}


\section{Introduction}

%\IEEEPARstart
Enhanced IP (EnIP), provides a model for increasing the
life of IPv4 by integrating EnIP into the current
IPv4 infrastructure.  One powerful feature of EnIP is that it
does not require updates to DHCP, ARP, and existing IPv4 routing
protocols.  EnIP packets transit the Internet as IPv4 packets
carrying additional address bits and state in an IP option, so
there is no need for a routing table update as is the case in IPv6.
However, EnIP's use of an IP option requires minor additional logic to
be added to fast path implementations on routers\cite{router01}.

\section{Outline}
In parallel to Enhanced IP research, a survey paper covering many aspects
of IPv4, IPv6, and Carrier Grade NAT was created.  The survey paper is used
to justify research on Enhanced IP.  The title of the survey paper is:
Survey of Deployed Methods for Increasing Internet Address Space.  For those not familiar with the transition
to IPv6 and the evolution of IPv4, it is recommended that the survey paper
be read prior to this paper.

Section III of this paper describes the design decision to use IP options in EnIP.
Section IV introduces the mechanics of EnIP, followed by section V which
discusses the integration plan for EnIP.  
The final section includes closing remarks with emphasis on some areas
for further research.

%\include makes a pagebreak
%\input{transitions}  


\input{ip_options}    
\input{eip_basics}     %bill
\input{eip_transition} %steve
\input{conclusion}	

\bibliographystyle{plain}
\bibliography{ourrefs,rfc}

\end{document}
